---
title: "R HW 1"
author: "Gabe Mancino"
date: "8/24/2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In class, we considered the case of rolling two dice, one red and one white.  Let $Y\equiv$ the minimum of the two rolls. 

#### 1. 
Define the probability mass function (pmf) in tabular format.  The table is started for you, below.  Fill in the appropriate entries for $p(y)$.

Probability mass function for $Y:=$ minimum of rolling one white and one red dice:

```{r}
y <- c(1:6)
py <- c(11/36, 9/36, 7/36, 5/36, 3/36,1/36)
dd <- data.frame(y = y, probs = py)
```
$y$  |  $p(y)$
---  |  ---
1    |  11/36
2    |  9/36
3    |  7/36
4    |  5/36
5    |  3/36
6    |  1/36

#### 2.
Graph this pmf.

```{r}
library(ggplot2)
dd <- data.frame(y = y, probs = py)
ggplot(aes(x = y, y = probs), data = dd) + geom_bar(stat = 'identity') + ylab('P(Y=y)') + ggtitle('pmf of Y')
```

#### 3.
Find $\mu \equiv E(Y)$.

$E(Y)=$
```{r}
ey <- sum(y * py)
ey
```

#### 4.
Find $\sigma^2 \equiv Var(Y)$.

$Var(Y)=$
```{r}
vy <- sum(y**2 * py) - ey**2
vy
```

#### 5.
Find $M_Y(t)$, and create a plot of $M_Y(t)$ as a function of $t$ over the interval $[-1,1]$.  (Hint: first, write a function `m(t)` that takes as argument `t` and returns $M_Y(t)$, for any input $t$.  Then, create a sequence of $t-$values over $[1,1]$ with the `seq()` function, and use your function to evaluate $M_Y(t)$ for each value in the sequence.  Then plot this sequence with $t$ on the x-axis and $M_Y(t)$ on the y-axis.  The function `exp(t)` can be used to evaluate $e^t$.)

```{r}
m.t <- function(t) {
  y <- c(1:6)
  py <- c(11/36, 9/36, 7/36, 5/36, 3/36, 1/36)
  e.t <- exp(y * t)
  f.p <- e.t * py
  f.t <- sum(f.p)
  return(f.t)
}
a.t <- seq(-1,1, by = 0.01)
out <- sapply(a.t, m.t)
df <- data.frame(a.t, out)
ggplot(aes(x = a.t, y = out), data = df) + geom_line(stat = 'identity') + ylab(expression(paste(M[y] (t)))) + xlab('Time') + ggtitle('mgf of Y')
```

#### 6.
Now write an `R` function `one.Y()` that simulates a single realization of $Y$.  Using this function and `replicate()`, generate 1000 realizations of $Y$.  Graph the empirical pmf, and verify that is is similar to the theoretical one above.

```{r}
one.Y <- function() {
  possible.rolls <- c(1:6)
  two.rolls <- sample(possible.rolls, 2, replace = TRUE)
  total <- min(two.rolls)
  return(total)
}
many.Y <- replicate(1000, one.Y())
df <- data.frame(x = as.factor(many.Y))
ggplot(aes(x = as.factor(many.Y)), data = df) + geom_bar(aes(y = (..count..)/(sum(..count..)))) + ylab('Observed Proportion') + xlab('y') + ggtitle('Empirical pmf of Y')
```

#### 7.
Find the empirical mean $\bar Y$ and variance $S^2$ (use `mean()` and `var()`).  How do they compare to $\mu$ and $\sigma^2$?

```{r}
mean(many.Y)
var(many.Y)
```
Clearly, the observed mean and the observed variance are "close" to $\mu$ and $\sigma^2$. 