---
title: "R Homework 1"
author: "Gabe Mancino"
date: "1/20/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let $Y$ represent the actual amount of coca-cola in 12-ounce bottles coming off a production line, with $Y \sim N(12, \sigma^2_Y = 0.01)$.  A manufacturing company claims to make filling machines that are more precise.  Specifically, with $X$ representing the actual amount of coke in 12-ounce bottles filled by the new machine, the company claims that $X\sim N(12, \sigma^2_X = 0.005)$.

The manager at the coke factory decides to test this claim with a sample $Y_1,...,Y_{20}$ taken i.i.d. from the existing machines, but a larger sample $X_1,...,X_{50}$ from the new machine.  


1. Simulate 1000 realizations of an appropriately scaled function of the sample variances, $s^2_X$ and $s^2_Y$, that follows an $F$ distribution.  Plot a histogram of these realizations, and superimpose the appropriate theoretical $F$ distribution. Verify that the theoretical distribution well-approximates the simulated realizations.

**Solution.**

Consider the following $R$ code.
```{r, tidy = TRUE, tidy.opts=list(width.cutoff=60)}
library(ggplot2)
one.samp <- function(){
  samp.x <- rnorm(20, mean = 12, sd = sqrt(0.005))
  samp.y <- rnorm(50, mean = 12, sd = sqrt(0.01))
  s2.x <- var(samp.x)/0.005
  s2.y <- var(samp.y)/0.01
  output <- s2.x/s2.y
  return(output)
}
many.samp <- replicate(1000, one.samp())
data.f <- data.frame(many.samp)
ggplot(aes(x = data.f), data = data.f) + geom_histogram(aes(y = ..density..), bins = 75, col = 'black') + xlim(0, 5) + xlab('X') + ylab('Density') + ggtitle('Simulated and Theoretical F-Distribution') + stat_function(fun = df, geom = 'line', col = 'red', args = list(df1 = 49, df2 = 19))

```
Thus we can see that the simulated F-distribution well approximates the theoretical F-distribution.

$\hfill\square$


2.  Specify the form of a 95% confidence interval for $\sigma^2_Y/\sigma^2_X$.  

**Solution.**

To derive the 95% confidence interval for $\frac{\sigma^2_Y}{\sigma^2_X},$ we can note that our above graph is for $F\sim\frac{s^2_X/\sigma^2_X}{s^2_Y/\sigma^2_Y}$, thus a 95% confidence interval is of the form, $P(Q_{0.025}\le F\le Q_{0.975}).$ Then some algebra yields,
\[
P(Q_{0.025}\le\frac{s^2_X/\sigma^2_X}{s^2_Y/\sigma^2_Y}\le Q_{0.975}) = P(\frac{s^2_Y}{s^2_X}Q_{0.025}\le\frac{\sigma^2_Y}{\sigma^2_X}\le\frac{s^2_Y}{s^2_X}Q_{0.975}).
\]

$\hfill\square$

\newpage
3.  Using the confidence interval you just derived, simulate 1000 realizations of these confidence intervals, assuming the manufacturing company's claims are correct.  Verify that the confidence intervals have approximately correct coverage of the true value of $\sigma^2_Y/\sigma^2_X$.  

**Solution.**

Consider the following $R$ code.
```{r, tidy = TRUE, tidy.opts=list(width.cutoff=60)}
#Generate one confidence interval with the following code:
get.one.ci <- function(){
  one.samp.y <- rnorm(20, mean = 12, sd = sqrt(0.01))
  one.samp.x <- rnorm(50, mean = 12, sd = sqrt(0.005))
  var.y <- var(one.samp.y)
  var.x <- var(one.samp.x)
  rat <- var.y / var.x
  lower <- rat * qf(0.025, df1 = 49, df2 = 19)
  upper <- rat * qf(0.975, df1 = 49, df2 = 19)
  ci <- c(lower, upper)
  return(ci)
}
#Create a function that tells us in our confidence interval over laps the true ratio of variances
covers <- function(ci){
  covers <- ifelse(ci[1] < 2 & 2 < ci[2], 'Yes', 'No')
  return(covers)
}
#Generate 1000 random confidence intervals
many.ci <- replicate(1000, get.one.ci())
data.ci <- data.frame(t(many.ci))
#Apply our cover function to each interval
data.ci$Coverage <- apply(data.ci, 1, covers)
data.ci$Sample <- 1:nrow(data.ci)
#Print out the percentages of 'Yes' and 'No'
table(data.ci$Coverage)/1000
```
Thus we can see that the confidence intervals have approximate the correct coverage.

$\hfill\square$


